CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-10-01 codecheck-paper.tex
small update to principles, editorial changes

The workflow and variations just outlined below reflect our current
views on how a codecheck should be performed. They are not set in
stone, but we do believe the following core principles underpin our
CODECHECK:

@@

\textbf{3. Credit is given to codecheckers.} The value of contributing a
CODECHECK is comparable to providing a scientific peer review. Indiviual
variations aside, it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e., with a DOI; this is done within the community codecheck
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases,
such as ORCID \footnote{\href{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}}
or Publons.

\textbf{4. Workflows must be auditable.} This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Tode could be executed at least once without critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
A codechecker does \emph{not} conduct a code review.
% more content from the website, may fit better elsewhere:
% The CODECHECK is not automated on purpose: automation may (a) lead to people gaming the system, (b) hide details that eventually decrease level of certainty that a codechecker has in their assessment, and (c) reduce the understandability of instructions in the long term, which is more important than short term ease of use (see Principle 2).

@@

\section*{Register}\label{register}
% count certificates: 
% curl https://raw.githubusercontent.com/codecheckers/register/master/register.csv | wc -l
% minus 1 for Hathway-Goodman (no empty last line, so no -1 for the headline needed)

@@

At the completion of each CODECHECK, a certificate is written that
states which outputs, i.e., numbers, figures and tables, from the 
original article could be reproduced by the CODECHECKER.
This certificate is made openly
available so that all readers of the article can see which elements
have been reproduced.  The certificate also includes links to the code
and data used by the codechecker, allowing others to build on the
work.
The format of the certificates has evolved slightly during the
project, as we have learnt to automate different aspects of the
certification.

@@

Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
lists the outputs that were generated from the codecheck (this is
termed the MANIFEST.)  For each output, there is a comment stating
which figure/table it should be compared to in the original paper,
along with the size of the file.  On page 3 of the certificate,
Figure~\ref{fig:cert}E shows some more detailed notes from the
codechecker, in this case indicating what steps were needed to
initialize the environment and run the code.  This particular example
was quite long, taking about 17 hours to run.  Finally, on page 4 of
the certificate, we see the first output that was generated by the
codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
figure 4 of REF.  Subsequent pages of the certificate show other
outputs (not shown here).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
  \caption{Annotated certificate 2020-012 (first four pages only).}
  \label{fig:cert}
\end{figure}

>>>>

The example workflow and variations just outlined above reflect our current
views on how a codecheck should be performed. They are not set in
stone, but we do believe the following core principles underpin our
CODECHECK:

@@

\textbf{3. Credit is given to codecheckers.} The value of contributing a
CODECHECK is comparable to providing a scientific peer review. Indiviual
variations aside, it may also require a similar amount of time. Therefore,
the codechecker's activity should be publicly recorded and mentioned in the
published paper.
The public record can be realised by publishing the certificate in a 
citable form (i.e., with a DOI; this is done within the community codecheck
process), by listing codecheckers on the journal's website or, ideally, by
publishing the checks alongside peer review activities in public databases.

\textbf{4. Workflows must be auditable.} This requires that the codechecker 
has at least enough material to validate the workflow outputs submitted by 
the authors. Stark~\cite{stark_before_2018} calls this `preproducibility'.
Code could be executed at least once without critical errors or
warnings using the provided instructions, and, therefore, all code and data 
was given and could be investigated more deeply or extended in the future.
Ideally, this is a "one~click" step, but achieving this requires particular 
skills. It is also hard to create a sufficient level of documentation for 
third parties. That is why the possibility to communicate is more important
than perfected workflows, and why a CODECHECK is not automated.
A codechecker does \emph{not} conduct a code review.
% more content from the website, may fit better elsewhere:
% The CODECHECK is not automated on purpose: automation may (a) lead to people gaming the system, (b) hide details that eventually decrease level of certainty that a codechecker has in their assessment, and (c) reduce the understandability of instructions in the long term, which is more important than short term ease of use (see Principle 2).

@@

\section*{Register}\label{register}
% count certificates: 
% curl https://codecheck.org.uk/register/stats.json
% -1 for Hathway-Goodman (no empty last line, so no -1 for the column headers needed)

@@

At the completion of each CODECHECK, a certificate is written that
states which outputs, i.e., numbers, figures and tables, from the 
original article could be reproduced by the codechecker.
This certificate is made openly
available so that all readers of the article can see which elements
have been reproduced.  The certificate also includes links to the code
and data used by the codechecker, allowing others to build on the
work.
The format of the certificates has evolved slightly during the
project, as we have learnt to automate different aspects of the
certification.

@@

Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
lists the outputs that were generated from the codecheck (this is
termed the MANIFEST).  For each output, there is a comment stating
which figure/table it should be compared to in the original paper,
along with the size of the file.  On page 3 of the certificate,
Figure~\ref{fig:cert}E shows some more detailed notes from the
codechecker, in this case indicating what steps were needed to
initialize the environment and run the code.  This particular example
was quite long, taking about 17 hours to run.  Finally, on page 4 of
the certificate, we see the first output that was generated by the
codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
figure~4 of \cite{Unwin2020}.  Subsequent pages of the certificate show other
outputs (not shown here).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/annotate-cert-crop.pdf}
  \caption{Annotated certificate 2020-012 \cite{cert-2020-012} (first four pages only).}
  \label{fig:cert}
\end{figure}
